---
title: "additional charts"
output:
  html_document:
    theme: paper
    toc: true
    toc_float:
      collapsed: false
---

<!-- Setting default chunk options -->
```{r chunkOptions, echo=FALSE, include = FALSE}
#knitr::opts_chunk$set(echo = TRUE)
# setting global R chunk options (https://yihui.name/knitr/options/#chunk_options)
  knitr::opts_chunk$set(dev='svglite', # svg, png,...
                        fig.ext = ".svg",
                        #fig.asp = .8 # default aspect ratio
                        fig.width = 12,
                        fig.height = 6,
                        dpi=100
                        )
```  


```{r loading_packages, echo=FALSE, include = FALSE}

  #Loading required packages
  # install.packages("remotes")
  #remotes::install_github("coolbutuseless/ggpattern")
  #install.packages('rsvg')
  #remotes::install_github('coolbutuseless/ggsvg')
  packagesList <- c("quitte","ggplot2","geomtextpath","ggpattern","tidyr","grid","stringr","gridExtra","ggrepel","kableExtra","dplyr","remind2","rsvg","ggsvg","plotly")
  packages <- suppressWarnings(suppressMessages(lapply(packagesList, function(x){ if(!require(x, character.only = T, quietly = T)){ install.packages(x); return(paste0("installed missing package: ", x)) } else paste0("required package is installed: ", x) } )))

```


<!-- Preprocess: prepare and load the data for use in the charts and markdown -->

```{r preprocess, echo=FALSE, include = FALSE}

#Options
forceUpdateData = FALSE
loadHist = TRUE

# data and output folders
folders <- file.info(list.files(path="./data", full.names = T))
dataFolder <- folders %>% tibble::rownames_to_column(var = "path") %>% arrange(desc(path)) %>% slice(1) %>% pull(path)

source("./R/preprocess.R") 

outFolder <- paste0("./output/",gsub("./data/","",dataFolder))
outChartsFolder <- paste0(outFolder,"/report_additionalCharts")
dir.create(paste0(outChartsFolder, "/png"), recursive = TRUE, showWarnings = FALSE)
dir.create(paste0(outChartsFolder, "/svg"), recursive = TRUE, showWarnings = FALSE)

```

<!-- units conversion -->

```{r unitsConversion, echo=FALSE, include = FALSE}

# conversion factors
  USS2017_2_EURO2020 <- 0.9502 # 1 USD 2017 = 0.9502 EURO 2020
  
  USS2017_2_EURO2025 <- 1.212 # 1 USD 2017 = 1.212 EURO 2020 (1 US$2017 = 1.30 * 0.932642 €2025 = 1.212 €2025)
  # exchange rate = average of February 28, March 31, April 11, 2025 (0.960138 +  0.925734 + 0.912056)/3 = 0.932642 https://www.ofx.com/en-au/forex-news/historical-exchange-rates/yearly-average-rates/
  # CPI -> 1 $2017 = 1.30$2025 -> https://www.in2013dollars.com/us/inflation/2017?amount=1
  
  currencyConversion <- USS2017_2_EURO2025

```

<!-- Initializing charts object -->
```{r chart_initialization, echo=FALSE, include = FALSE}

mainScen <- "Nzero_57_bio7p5"
tgt2030Scen <- c(55,57,59)

#loading theme and colors
source("./R/aesthetics.R") 

g <- NULL

```


# Targets debug {.tabset}

## Reduction Matrix


```{r data_emissionReductionMatrix, echo=FALSE, include = FALSE}

## Reduction Matrix
reductionMatrix <- df %>% 
  filter(region == "EU27", 
         period == 2040, 
         tgt2030 == "57",
         !grepl("eedEff|ff55Eff|RpEUEff",scenario)) %>% 
  calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>%
  mutate(reduction = round((1 - (value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)))*100 ,2)) %>%
  select(bioLim,lim,reduction) %>%
  pivot_wider(names_from=lim,values_from=reduction) %>%
  mutate(bioLim = factor(c("BioLim4","BioLim7.5","BioLim12","BioLim20")))

if("limVRE" %in% colnames(reductionMatrix)){
  colors <- spec_color(c(reductionMatrix$default, reductionMatrix$unlimCCS,reductionMatrix$limCCS,reductionMatrix$limH2,reductionMatrix$limVRE,reductionMatrix$limVRE_limInt, reductionMatrix$limVRE_limInt_int), end = 0.9, option = "A", direction = -1)

redMatrix <- reductionMatrix %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = T) %>%
  column_spec(2,color = colors[1:3]) %>%
  column_spec(3,color = colors[4:6]) %>%
  column_spec(4,color = colors[7:9]) %>%
  column_spec(5,color = colors[10:12])
} else {
  colors <- spec_color(c(reductionMatrix$default,reductionMatrix$limCCS,reductionMatrix$limH2), end = 0.9, option = "A", direction = -1)

redMatrix <- reductionMatrix %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = T) %>%
  column_spec(2,color = colors[1:3]) %>%
  column_spec(3,color = colors[4:6]) %>%
  column_spec(4,color = colors[7:9])
}

redMatrix2 <- df %>%
    filter(region == "EU27", period == 2040, tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>%
    mutate(reduction = value / (tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)),
           tgtAchieved = (1-reduction)*100,
           diff = as.numeric(as.character(tgt2030)) - ((1-reduction)*100)) %>%
  mutate(text = paste0(round(tgtAchieved,2), "%"))

 bioLim_levels <- c(
   "BioLim4"="4 EJ",
   "BioLim7.5"="7.5 EJ", 
   "BioLim12"="12 EJ", 
   "BioLim20"="20 EJ")

lim_levels <- c(
  "default"="default",
  "unlimCCS"="unlimited CCS",
  "limCCS"="limited CCS",
  "limH2"="limited H2",
  "limVRE"="limited VRE",
  "limVRE_limInt"="lim VRE&Integ._alt",
	"limVRE_strgLimInt"="lim VRE&Integ.")

  tgt_levels <- c(
    "55"="55%",
    "57"="57%",
    "59"="59%")
  
  eff_levels <- c(
    "reference"="reference", 
    "EED 2018 Eff"="EED 2018", 
    "FitFor55 eff"="FitFor55", 
    "RePowerEU eff"="RePowerEU")

redMatrix2 <- redMatrix2 %>%
  mutate(
    tgt2030   = factor(tgt_levels[as.character(tgt2030)], levels = tgt_levels),
    efficiency= factor(eff_levels[as.character(efficiency)], levels = eff_levels),
    bioLim    = factor(bioLim_levels[as.character(bioLim)], levels = bioLim_levels),
    lim       = factor(lim_levels[as.character(lim)], levels = rev(lim_levels))
  )

g[["reduction matrix"]][["ggplot"]] <- ggplot(redMatrix2, aes(x=bioLim, y=lim, fill= tgtAchieved, text = text)) + 
  geom_tile() +
  facet_grid(tgt2030 ~ efficiency, scales = "free") +
  scale_fill_distiller(palette = "Blues", direction = 1, breaks = seq(floor(min(redMatrix2$tgtAchieved)), ceiling(max(redMatrix2$tgtAchieved)),2)) +
  theme_doubleColumn() +
  theme(strip.clip = "off") +
  #guides(fill = guide_colorbar(barwidth = 20, barheight = 1)) +
  theme(legend.position="right") +
  theme(legend.title=element_blank())
  
g[["reduction matrix"]][["plotly"]] <- ggplotly(g[["reduction matrix"]][["ggplot"]],tooltip = c("text")) %>% layout(autosize = TRUE)  

ggsave(paste0(outChartsFolder,"/svg/Emission Reduction 2040.svg"), g[["reduction matrix"]][["ggplot"]] , device="svg", width = 10, height = 6, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/Emission Reduction 2040.png"), g[["reduction matrix"]][["ggplot"]] , device="png", width = 10, height = 6, dpi=300, units = "in")

```

```{r, echo=FALSE, fig.width=10, fig.height=6}

g[["reduction matrix"]][["plotly"]]


```

- Reduction Matrix for the 2030 57% target

```{r, echo=FALSE}

redMatrix

```


## non-achieved 2030

<!-- Debug non-achieved targets -->
```{r, echo=FALSE, include = FALSE}


plotData <- left_join(
  df %>%
    filter(region == "EU27", period == 2030, tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>%
    mutate(reduction = value / (tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)),
           tgtAchieved = (1-reduction)*100,
           diff = as.numeric(as.character(tgt2030)) - ((1-reduction)*100)) %>%
    group_by(tgt2030) %>% 
    mutate(min = min(value),
           max = max(value)),
  df %>%
    filter(region == "EU27", period == 2030, tgt2030 %in% tgt2030Scen, variable == "Price|Carbon") %>%
    mutate(price_2030 = currencyConversion*value) %>%
    select(scenario, price_2030),
  by = "scenario") %>%
  mutate(text = paste0("Target Achieved: ", tgtAchieved, " % reduction", "\n2030 price: ", round(price_2030,2), " €/t CO2", "\nScenario: ", scenario, "\nTarget 2030: ", tgt2030, "%", "\nEfficiency: ", efficiency, "\nBiomass Limit: ", bioLim, "\nSensititivy limit: ", lim))

g[["2030 targets"]][["ggplot"]] <-  ggplot(plotData,aes(x=period,y=tgtAchieved)) +
  geom_point(aes(shape = lim, fill = bioLim, color = efficiency, text = text), size = 3) +
  geom_segment(aes(x=2029, xend=2031, y=as.numeric(as.character(tgt2030)), yend=as.numeric(as.character(tgt2030))), linetype="dashed", color="black") +
  #geom_errorbar(aes(ymin=min, ymax=max), colour="black", alpha=0.9, linewidth=1) +
  #geom_boxplot(outlier.shape = NA) +
  facet_grid(~ tgt2030) +
  theme_doubleColumn() +
  scale_fill_manual(values = c("BioLim4"="#fedc97", "BioLim7.5"="#b5b682", "BioLim12"="#7c9885", "BioLim20"="#28666e"), guide = guide_legend(nrow =2, byrow = F, override.aes = list(shape = 21, size = 6, colour = "transparent"), order = 3)) +
  scale_shape_manual(values = c("default"=21,"unlimCCS"=24,"limCCS"=25,"limH2"=23,"limVRE"=22,"limVRE_limInt"=12,"limVRE_strgLimInt"=14), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 6))) +
  scale_color_manual(values = c("reference"="#ae0000", "EED 2018 Eff"="#db4646", "FitFor55 eff"="#0090ab", "RePowerEU eff"="#095786"), guide = guide_legend(nrow =2, byrow = F, order = 2, override.aes = list(shape = 1, linewidth = 1.5))) + 
  scale_x_continuous(breaks=2030) +
  scale_y_continuous(breaks=seq(53,61,1)) +
  theme(legend.position="none")
  
g[["2030 targets"]][["plotly"]] <- ggplotly(g[["2030 targets"]][["ggplot"]],tooltip = c("text")) %>% layout(autosize = TRUE)

ggsave(paste0(outChartsFolder,"/svg/Emission Reduction 2030.svg"), g[["2030 targets"]][["ggplot"]], device="svg", width = 8, height = 4, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/Emission Reduction 2030.png"), g[["2030 targets"]][["ggplot"]], device="png", width = 8, height = 4, dpi=300, units = "in")

## Function to extract legend
g_legend <- function(a.gplot){ 
    tmp <- ggplot_gtable(ggplot_build(a.gplot)) 
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box") 
    legend <- tmp$grobs[[leg]] 
    legend
} 

legend <- g_legend(
  g[["2030 targets"]][["ggplot"]] +
  theme(legend.position="bottom") +  
  theme(legend.title=element_blank(),
        legend.margin=margin(c(0,0,0,0))) +
  theme(plot.margin=margin(0,0,0,0))
) 

nonAchieved2030 <- left_join(
  df %>%
    filter(region == "EU27", period == 2030, tgt2030 %in% tgt2030Scen) %>%
    calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>%
    mutate(reduction = value / (tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)),
           tgtAchieved = (1-reduction)*100,
           diff = as.numeric(as.character(tgt2030)) - ((1-reduction)*100)) %>%
    group_by(tgt2030) %>% 
    mutate(min = min(value),
           max = max(value)),
  df %>%
    filter(region == "EU27", period == 2030, tgt2030 %in% tgt2030Scen, variable == "Price|Carbon") %>%
    mutate(price_2030 = currencyConversion*value) %>%
    select(scenario, price_2030),
  by = "scenario")

underAchieved2030 <- nonAchieved2030 %>%
  filter(diff >= 0.5) %>%
  select(scenario, price_2030, tgt2030, tgtAchieved, diff) %>%
  arrange(desc(abs(diff)))

overAchieved2030 <- nonAchieved2030 %>%
  filter(diff <= 0.5) %>%
  select(scenario, price_2030, tgt2030, tgtAchieved, diff) %>%
  arrange(desc(abs(diff)))

rerun <- overAchieved2030 %>% filter(tgtAchieved >= 59.5)
  
#View(nonAchieved2030)

```


```{r, echo=FALSE, fig.width=8, fig.height=4}

g[["2030 targets"]][["plotly"]]

```

```{r, echo=FALSE, fig.width=8, fig.height=2}

grid.newpage()
grid.draw(legend)

```


```{r, echo=FALSE}

nonAchieved2030 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```


## non-achieved 2050

<!-- Debug non-achieved targets -->
```{r, echo=FALSE, include = FALSE}


plotData <- df %>%
  filter(region == "EU27", period == 2050, tgt2030 %in% tgt2030Scen) %>%
  calc_addVariable( "`Emissions with LULUCF and with international transport`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr", only.new=T) %>%
  mutate(reduction = value / (tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)),
         tgtAchieved = (1-reduction)*100) %>% 
  mutate(min = min(value),
         max = max(value))

g[["2050 targets"]][["ggplot"]] <-  ggplot(plotData,aes(x=period,y=tgtAchieved)) +
  geom_point(aes(shape = lim, fill = bioLim, color = efficiency), size = 3) +
  geom_segment(aes(x=2049, xend=2051, y=100, yend=100), linetype="dashed", color="black") +
  theme_doubleColumn() +
  scale_fill_manual(values = c("BioLim4"="#fedc97", "BioLim7.5"="#b5b682", "BioLim12"="#7c9885", "BioLim20"="#28666e"), guide = guide_legend(nrow =2, byrow = F, override.aes = list(shape = 21, size = 6, colour = "transparent"), order = 3)) +
  scale_shape_manual(values = c("default"=21,"unlimCCS"=24,"limCCS"=25,"limH2"=23,"limVRE"=22,"limVRE_limInt"=12,"limVRE_strgLimInt"=14), guide = guide_legend(nrow =2, byrow = T, order = 4, override.aes = list(size = 6))) +
  scale_color_manual(values = c("reference"="#ae0000", "EED 2018 Eff"="#db4646", "FitFor55 eff"="#0090ab", "RePowerEU eff"="#095786"), guide = guide_legend(nrow =2, byrow = F, order = 2, override.aes = list(shape = 1, linewidth = 1.5))) + 
  scale_x_continuous(breaks=2050) +
  scale_y_continuous(breaks=c(99.5,100,100.5)) +
  expand_limits(y=c(99.5,100.5)) +
  theme(legend.position="none")

  g[["2050 targets"]][["plotly"]] <- ggplotly(g[["2050 targets"]][["ggplot"]]) %>% layout(autosize = TRUE)

  
ggsave(paste0(outChartsFolder,"/svg/Emission Reduction 2050.svg"), g[["2050 targets"]][["ggplot"]], device="svg", width = 4, height = 4, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/Emission Reduction 2050.png"), g[["2050 targets"]][["ggplot"]], device="png", width = 4, height = 4, dpi=300, units = "in")

nonAchieved2050 <- df %>%
  filter(region == "EU27", period == 2050, tgt2030 %in% tgt2030Scen) %>%
  calc_addVariable( "`Emissions with LULUCF and with international transport`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers`" , units = "Mt CO2/yr", only.new=T) %>%
  mutate(reduction = value / (tgt %>% filter(target == "Emissions with LULUCF and with international transport") %>% pull(reference)),
         tgtAchieved = (1-reduction)*100) %>%
  filter(tgtAchieved >= 100.3 | tgtAchieved <= 99.7) %>%
  select(scenario, reduction, tgtAchieved) %>%
  arrange(desc(abs(reduction)))

#View(nonAchieved2050)

```


```{r, echo=FALSE, fig.width=4, fig.height=4}

g[["2050 targets"]][["plotly"]]

```


```{r, echo=FALSE, fig.width=8, fig.height=2}

grid.newpage()
grid.draw(legend)

```

```{r, echo=FALSE}

nonAchieved2050 %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```




# Fuel Shares {.tabset}

```{r, echo=FALSE, include=FALSE}

vars <- list(
  "Liquids" = list(
    "Transport" = c(
      "FE|Transport|Liquids|Biomass",
      "FE|Transport|Liquids|Hydrogen",
      "FE|Transport|Liquids|Fossil"
    ),
    "Buildings" = c(
      "FE|Buildings|Liquids|Biomass",
      "FE|Buildings|Liquids|Hydrogen",
      "FE|Buildings|Liquids|Fossil"
    ),
    "Industry" = c(
      "FE|Industry|Liquids|Biomass",
      "FE|Industry|Liquids|Hydrogen",
      "FE|Industry|Liquids|Fossil"
    )
  ),
  "Gases" = list(
    "Transport" = c(
      "FE|Transport|Gases|Biomass",
      "FE|Transport|Gases|Hydrogen",
      "FE|Transport|Gases|Fossil"
    ),
    "Buildings" = c(
      "FE|Buildings|Gases|Biomass",
      "FE|Buildings|Gases|Hydrogen",
      "FE|Buildings|Gases|Fossil"
    ),
    "Industry" = c(
      "FE|Industry|Gases|Biomass",
      "FE|Industry|Gases|Hydrogen",
      "FE|Industry|Gases|Fossil"
    )
  ),
  "Solids" = list(
    "Buildings" = c(
      "FE|Buildings|Solids|Biomass",
      "FE|Buildings|Solids|Fossil"
    ),
    "Industry" = c(
      "FE|Industry|Solids|Biomass",
      "FE|Industry|Solids|Fossil"
    )
  )
  )

plotColor <- c(
    "Biomass"  = "#597445",
    "Hydrogen" = "#F96666",
    "Fossil"   = "#bbbbbb"
  )


#tgt2050 tgt2030 efficiency bioLim lim  

fullPlotData <- df %>%
  filter(region == "EU27", period %in% c(2010:2050), variable %in% unlist(vars)) %>%
  mutate(sector = factor(unlist(regmatches(variable, gregexpr('Transport|Industry|Buildings', variable))),levels=c("Industry", "Buildings", "Transport")),
         carrier= factor(unlist(regmatches(variable, gregexpr('Liquids|Gases|Solids', variable))),levels=c("Liquids", "Gases", "Solids")),
         type   = factor(unlist(regmatches(variable, gregexpr('Biomass|Hydrogen|Fossil', variable))),levels=c("Biomass", "Fossil", "Hydrogen")))

plotData <- fullPlotData %>%
  filter(tgt2030 == 57, 
         efficiency == "reference", 
         #bioLim == "BioLim7.5", 
         lim == "default")

g[["Fuel shares"]][["BioLim"]] <-  ggplot(plotData,aes(x=period,y=value)) +
    geom_bar(aes(fill=type),stat='identity', position="fill") +
    facet_grid(carrier~bioLim+sector) +
    theme_doubleColumn() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = plotColor) + 
    theme(plot.background = element_rect(fill = 'white', colour = 'white'))

ggsave(paste0(outChartsFolder,"/svg/Fuel shares - tgt57,effRef,limDef - BioLim.svg"), g[["Fuel shares"]][["BioLim"]], device="svg", width = 12, height = 8, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/Fuel shares - tgt57,effRef,limDef - BioLim.png"), g[["Fuel shares"]][["BioLim"]], device="png", width = 12, height = 8, dpi=300, units = "in")

plotData <- fullPlotData %>%
  filter(tgt2030 == 57, 
         #efficiency == "reference", 
         bioLim == "BioLim7.5", 
         lim == "default")

g[["Fuel shares"]][["efficiency"]] <-  ggplot(plotData,aes(x=period,y=value)) +
    geom_bar(aes(fill=type),stat='identity', position="fill") +
    facet_grid(carrier~efficiency+sector) +
    theme_doubleColumn() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = plotColor) + 
    theme(plot.background = element_rect(fill = 'white', colour = 'white'))

ggsave(paste0(outChartsFolder,"/svg/Fuel shares - tgt57,bio7p5,limDef - efficiency.svg"), g[["Fuel shares"]][["efficiency"]], device="svg", width = 12, height = 8, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/Fuel shares - tgt57,bio7p5,limDef - efficiency.png"), g[["Fuel shares"]][["efficiency"]], device="png", width = 12, height = 8, dpi=300, units = "in")

plotData <- fullPlotData %>%
  filter(#tgt2030 == 57, 
         efficiency == "reference", 
         bioLim == "BioLim7.5", 
         lim == "default")

g[["Fuel shares"]][["tgt2030"]] <-  ggplot(plotData,aes(x=period,y=value)) +
    geom_bar(aes(fill=type),stat='identity', position="fill") +
    facet_grid(carrier~tgt2030+sector) +
    theme_doubleColumn() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = plotColor) + 
    theme(plot.background = element_rect(fill = 'white', colour = 'white'))

ggsave(paste0(outChartsFolder,"/svg/Fuel shares - effRef,bio7p5,limDef - tgt2030.svg"), g[["Fuel shares"]][["tgt2030"]], device="svg", width = 12, height = 8, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/Fuel shares - effRef,bio7p5,limDef - tgt2030.png"), g[["Fuel shares"]][["tgt2030"]], device="png", width = 12, height = 8, dpi=300, units = "in")

plotData <- fullPlotData %>%
  filter(tgt2030 == 57, 
         efficiency == "reference", 
         bioLim == "BioLim7.5"#, 
         #lim == "default"
         )

g[["Fuel shares"]][["lim"]] <-  ggplot(plotData,aes(x=period,y=value)) +
    geom_bar(aes(fill=type),stat='identity', position="fill") +
    facet_grid(carrier~lim+sector) +
    theme_doubleColumn() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = plotColor) + 
    theme(plot.background = element_rect(fill = 'white', colour = 'white'))

ggsave(paste0(outChartsFolder,"/svg/Fuel shares - tgt57,effRef,bio7p5 - lim.svg"), g[["Fuel shares"]][["lim"]], device="svg", width = 12, height = 8, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/Fuel shares - tgt57,effRef,bio7p5 - lim.png"), g[["Fuel shares"]][["lim"]], device="png", width = 12, height = 8, dpi=300, units = "in")

```


## tgt2030

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["Fuel shares"]][["tgt2030"]]

```

## Efficiency

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["Fuel shares"]][["efficiency"]]

```

## BioLim

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["Fuel shares"]][["BioLim"]]

```

## Lim

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["Fuel shares"]][["lim"]]

```





# Power sector Emissions


```{r, echo=FALSE, include=FALSE}

baseSize <- 24

theme_white <- function(){
  theme_minimal(base_size = baseSize) +
    theme(
      legend.position="bottom",
      legend.title = element_blank(),
      axis.title.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
      legend.key.height = unit(1, "cm"),
      legend.key.width = unit(1,"cm"),
      panel.background = element_rect(fill="#FFFFFF", color = NA),
      plot.background = element_rect(fill="#FFFFFF", color = NA),
    )
}

histEmber <- data.frame(period=seq(2000,2023,1),value=c(1079, 1084, 1122, 1166, 1157, 1156, 1170, 1188, 1128, 1039, 1048, 1054, 1033, 980, 917, 948, 934, 942, 879, 769, 660, 722, 750, 596))

mainScen <- "Nzero_57_bio7p5"

plotData <- df %>%
  filter(region == "EU27", period %in% c(2010:2050), variable == "Emi|CO2|Energy|Supply|Electricity w/ couple prod", scenario != "Npi") %>%
  group_by(period) %>%
  summarise_at(vars(value), list(
    Min = min,
    Mean = mean,
    Median = median,
    Max = max,
    Sd = sd,
    Q1=~quantile(., probs = 0.25),
    Q3=~quantile(., probs = 0.75)))

plotData2 <- df %>%
  filter(region == "EU27", period %in% c(2010:2050), variable == "Emi|CO2|Energy|Supply|Electricity w/ couple prod", scenario != "Npi") %>%
  filter(scenario == mainScen)

g[["power_sector_emissions"]] <- ggplot(plotData,aes(x=period,y=Median)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    #geom_line(linewidth=1,color="black") +
    geom_line(data = histEmber,aes(x=period,y=value), linewidth=1,color="black",linetype="dashed") +
    geom_line(data = plotData2,aes(x=period,y=value), linewidth=1,color="black") +
    geom_ribbon(aes(ymax = Q3, ymin = Q1), alpha = 0.5, fill = "#545454") +
    geom_ribbon(aes(ymax = Max, ymin = Min), alpha = 0.5, fill = "#545454") +
    #geom_line(aes(y=Min),linewidth=1,color="#545454",linetype = "dashed", alpha = 0.7) +
    #geom_line(aes(y=Max),linewidth=1,color="#545454",linetype = "dashed", alpha = 0.7) +
    #geom_line(data=h,linewidth=1,aes(linetype=Source)) +
    #geom_point() +
    theme_white() +
    ylab(expression(paste("Mt ", CO[2],"/yr")))


ggsave(paste0(outChartsFolder,"/svg/power_sector_emissions.svg"), g[["power_sector_emissions"]], device="svg", width = 12, height = 8, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/power_sector_emissions.png"), g[["power_sector_emissions"]], device="png", width = 12, height = 8, dpi=300, units = "in")


```

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["power_sector_emissions"]]

```


# VRE Analysis

```{r, echo=FALSE, include=FALSE}

# Define main scenario for comparison (fixing other dimensions)
# e.g. target 57, efficiency reference, bioLim 7.5
tgt_select <- "55"
eff_select <- "reference"
bio_select <- "BioLim4" # Using label from processed data if mapped, but filtering usually happens before full mapping or we use internal codes. 
# Checking previous chunks: bioLim labels are like "BioLim7.5"
# lim labels are "limited VRE", "lim VRE integration", etc. if mapped.
# However, usually easier to filter on IDs if possible, but loadData maps them.
# Let's check `df` structure from previous chunks.
# It seems `modifyData` maps them.
# modifyData uses:
# "limVRE" = "limited VRE"
# "limVREandIntegration" = "limVRE_limInt" -> Mapped to "lim VRE&Int" in paperCharts, but here in additionalCharts?
# Let's check filter logic.

# In this file, `df` is used.
# Let's reuse `plotData` pattern.

vars_vre <- c(
  "SE|Electricity|Solar|PV",
  "SE|Electricity|Wind|Onshore",
  "SE|Electricity|Wind|Offshore"
)

# Filter for VRE scenarios
vres_lims <- c("limVRE", "limVRE_limInt", "limVRE_strgLimInt")
# We need to match the actual levels in the dataframe.
# Based on the user's previous edits in paperCharts, the mapping was:
# "limVRE"="limited VRE",
# "limVRE_limInt"="lim VRE&Int",
# "limVRE_strgLimInt"="lim VRE&Int2"
#
# BUT, `report_additionalCharts.Rmd` might not have the same factors applied as `report_paperCharts.Rmd`.
# Let's inspect `redMatrix2` chunk in this file (lines 127+).
# It uses `lim` column.
#
# If the user wants to compare "scenarios containing with lim = limVRE, ...", they likely mean the scenario ID substring or the `lim` column.
# Let's assume the `lim` column has been populated by `loadData.R` with the keys from `limList` (values of `limList`).
# original keys: limVRE, limVRE2, limVRE3
# original values in limList: limVRE, limVRE_limInt, limVRE_strgLimInt
# So `df$lim` should contain "limVRE", "limVRE_limInt", "limVRE_strgLimInt".

plotDataVRE <- df %>%
  filter(region == "EU27", 
         period %in% c(2030, 2040, 2050),
         variable %in% vars_vre,
         lim %in% vres_lims) 

# Comparison Chart: Equivalent scenarios
# We fix efficiency and bioLim but show across different 2030 targets
compData <- plotDataVRE %>%
  filter(#tgt2030 == tgt_select, # Show all targets
         efficiency == eff_select,
         bioLim == bio_select)

g[["vre_comparison"]] <- ggplot(compData, aes(x=as.factor(period), y=value, fill=lim)) +
  geom_col(position = "dodge") +
  facet_grid(variable ~ tgt2030, scales = "free_y") +
  theme_doubleColumn() +
  labs(title = "VRE Capacities Comparison",
       subtitle = paste0("Eff: ", eff_select, ", Bio: ", bio_select),
       y = "GW", x = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(paste0(outChartsFolder,"/svg/VRE_Comparison.svg"), g[["vre_comparison"]], device="svg", width = 12, height = 8)
ggsave(paste0(outChartsFolder,"/png/VRE_Comparison.png"), g[["vre_comparison"]], device="png", width = 12, height = 8)


# Distribution Chart
# Comparing distribution across ALL scenarios that have these lim tags
g[["vre_distribution"]] <- ggplot(plotDataVRE, aes(x=lim, y=value, fill=lim)) +
  geom_boxplot() +
  #geom_violin(alpha=0.5) +
  facet_grid(period~variable, scales = "free_y") +
  theme_doubleColumn() +
  labs(title = "Distribution of VRE Capacities by Limitation Type",
       y = "GW") +
  theme(axis.text.x = element_blank()) # Label is in legend/x-axis tick might be redundant if color used

ggsave(paste0(outChartsFolder,"/svg/VRE_Distribution.svg"), g[["vre_distribution"]], device="svg", width = 12, height = 8)
ggsave(paste0(outChartsFolder,"/png/VRE_Distribution.png"), g[["vre_distribution"]], device="png", width = 12, height = 8)


```

## Comparison

```{r, echo=FALSE, fig.width=12, fig.height=8}
g[["vre_comparison"]]
```

## Distribution

```{r, echo=FALSE, fig.width=12, fig.height=8}
g[["vre_distribution"]]
```





# Milestones - ETL poster for the SAB review



concretely, he would like only 4 indicators: 

Emission reductions
BECCS and DACCS
Electricity share in FE
Wind & Solar share in Electricity (not yet part of the plot)


```{r chart_figure5, echo=FALSE}

g[["milestone"]] <- NULL

violin_scale <- 3 
violin_color = "#bbbbbb"
violin_boxPlotWidth <- 1.5   

#Emissions Reduction

plotData <- df %>%
  filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen) %>%
  calc_addVariable( "`Emissions with LULUCF and with intra-EU aviation`" = "`Emi|GHG|LULUCF national accounting` + `Emi|CO2|Energy|Demand|Transport|International Bunkers|Intra-region`" , units = "Mt CO2/yr", only.new=T) %>%
  mutate(value = round(1-(value / tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)) ,2)) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Emissions\nreductions") %>%
  group_by(period) %>%
  mutate(min = min(value),
        max = max(value))

h <- refEmi %>% filter(variable == "Emissions with LULUCF and with intra-EU aviation", region == "EU27", pollutant == "All greenhouse gases - (CO2 equivalent)") %>% ungroup() %>% select(region,variable,period,value)
emi1990 <- tgt %>% filter(target == "Emissions with LULUCF and with intra-EU aviation") %>% pull(reference)
reduction_h <- 1- (h %>% filter(period==2022) %>% pull(value) / emi1990)

target <- data.frame(period = 2030, x = 2025, xend= 2035, value = 0.55, label = "FitFor55")

g[["milestone"]][["Emissions reductions"]] <- ggplot(plotData %>% mutate(target = paste0(gsub("\n", " ", target)," (%)")), aes(x = period, y = value, group=period)) +
  geom_violin(fill="#bbbbbb", scale = violin_scale) + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=1.5) +
  geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  annotate("text", x = -Inf, y = 0.55, label = "FitFor55", angle=90, vjust = -0.5, hjust = 1, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_hline(yintercept=reduction_h, color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = Inf, y = reduction_h, label = "2022", angle=90, vjust = -0.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,1)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=0, b=20)))


# Novel CDR = atmospheric/biogenic CO2 (CDR) 

plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Carbon Management|atmospheric/biogenic CO2 - CDR") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "BECCS\nand\nDACCS") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value)) 

plotData <- plotData %>% mutate(target = paste0(gsub("\n", " ", target)," (Mt CO2/yr)"))

mainScenarioValue <- plotData %>% filter(scenario == mainScen)

g[["milestone"]][["Novel CDR"]] <- ggplot(plotData, aes(x = period, y = value, group=period)) +
  geom_violin(fill=violin_color, scale = violin_scale)  + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=violin_boxPlotWidth) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,300)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=0, b=20)))



# Direct Electrification

plotData <- df %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "Final Energy|w/o Bunkers|w/o Non-energy Use|Electricity Share") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Electricity\nShare in\nFinal Energy") %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value)) 

#estimation of historical share in 2022 for FE electricity / FE with non-energy use and with bunkers
# h <- (
#       2550461.10 *1000 / 8.76E+09 * 31.536  # https://ec.europa.eu/eurostat/databrowser/bookmark/12d54df4-7c35-47eb-a103-dd8994d26b80?lang=en
#      ) / (
#       addHist$FE_consumption %>% mutate(value = value/1000000) %>% filter(period==2022) %>% pull(value) + # FE|w/o Non-energy Use w/o Bunkers - EJ
#       df %>% filter(region == "EU27", period == 2020, scenario == "Npi", variable == "FE|Transport|Bunkers") %>% pull(value) +
#       df %>% filter(region == "EU27", period == 2020, scenario == "Npi", variable == "FE|Non-energy Use|Industry") %>% pull(value)
#      )

plotData <- plotData %>% mutate(target = paste0(gsub("\n", " ", target)," (%)"))

mainScenarioValue <- plotData %>% filter(scenario == mainScen)

g[["milestone"]][["Direct Electrification"]] <- ggplot(plotData, aes(x = period, y = value, group=period)) +
  geom_violin(fill=violin_color, scale = violin_scale)  + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=violin_boxPlotWidth) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,1)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=0, b=20))) 


# Wind & Solar share in Electricity

data <- df %>% calc_addVariable("`SE|wind and Solar share`" = "( (`SE|Electricity|Wind` + `SE|Electricity|Solar`) / `SE|Electricity`)", units = "%", only.new=T)

plotData <- data %>% filter(region == "EU27", period %in% c(2030,2040), tgt2030 %in% tgt2030Scen, variable == "SE|wind and Solar share") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Wind & Solar\nShare in\nElectricity") 

histValue_2023 <- hist %>% calc_addVariable("`SE|wind and Solar share`" = "( (`SE|Electricity|Wind` + `SE|Electricity|Solar`) / `SE|Electricity`)", units = "%", only.new=T) %>% filter(region == "EU27", model == "Ember", period == "2023")

plotData <- plotData %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

#fullPlotData <- rbind(fullPlotData,plotData) 

plotData <- plotData %>% mutate(target = paste0(gsub("\n", " ", target)," (%)"))
                                
mainScenarioValue <- plotData %>% filter(scenario == mainScen)

g[["milestone"]][["Wind & Solar Share in Electricity"]] <- ggplot(plotData, aes(x = period, y = value, group=period)) +
  geom_violin(fill=violin_color, scale = violin_scale)  + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=violin_boxPlotWidth) +
  #geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  #annotate("text", x = Inf, y = 0.4, label = "RED II revised", angle=90, vjust = -0.5, hjust = 0.075, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  #annotate("text", x = Inf, y = 0.45, label = "REPowerEU", angle=90, vjust = 1.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_hline(yintercept=histValue_2023 %>% pull(value), color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = Inf, y = histValue_2023 %>% pull(value), label = "2023", angle=90, vjust = -0.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  scale_x_reverse(breaks=c(2030,2040), labels = c(2030,2040), minor_breaks = NULL) +
  expand_limits(y=c(0,1)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=0, b=20)),
        #strip.text.x = element_text(size = textSize.double, hjust = 1, margin=margin(l=0, b=20)),
        strip.clip = "off") 

```


```{r, echo=FALSE, include=FALSE}

g[["milestones ETL poster"]] <- gridExtra::arrangeGrob(
  g[["milestone"]][["Emissions reductions"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Novel CDR"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Direct Electrification"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  g[["milestone"]][["Wind & Solar Share in Electricity"]], #+ theme(plot.margin=margin(l = 30, r = 10,t = 30)),
  ncol = 1
)

ggsave(paste0(outChartsFolder,"/svg/milestones ETL poster.svg"), g[["milestones ETL poster"]], device="svg", width = 4*1.5, height = 5*1.5, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/milestones ETL poster.png"), g[["milestones ETL poster"]], device="png", width = 4*1.5, height = 5*1.5, dpi=300, units = "in")



```


```{r, echo=FALSE, fig.width=8, fig.height=10}

grid.newpage()
grid.draw(g[["milestones ETL poster"]])

```


```{r, echo=FALSE, include=FALSE}


# Wind & Solar share in Electricity

data <- df %>% calc_addVariable("`SE|wind and Solar share`" = "( (`SE|Electricity|Wind` + `SE|Electricity|Solar`) / `SE|Electricity`)", units = "%", only.new=T)

plotData <- data %>% filter(region == "EU27", period %in% c(2030,2040, 2050), tgt2030 %in% tgt2030Scen, variable == "SE|wind and Solar share") %>% select(-variable) %>%
  filter(tgt2050 != "Npi") %>%
  mutate(target = "Wind & Solar\nShare in\nElectricity") 

histValue_2023 <- hist %>% calc_addVariable("`SE|wind and Solar share`" = "( (`SE|Electricity|Wind` + `SE|Electricity|Solar`) / `SE|Electricity`)", units = "%", only.new=T) %>% filter(region == "EU27", model == "Ember", period == "2023")

plotData <- plotData %>%
  group_by(period) %>% 
  mutate(min = min(value),
        max = max(value))

#fullPlotData <- rbind(fullPlotData,plotData) 

plotData <- plotData %>% mutate(target = paste0(gsub("\n", " ", target)," (%)"))
                                
mainScenarioValue <- plotData %>% filter(scenario == mainScen)

g[["milestone"]][["Wind & Solar Share in Electricity extended"]] <- ggplot(plotData, aes(x = period, y = value, group=period)) +
  geom_violin(fill=violin_color, scale = violin_scale)  + #, trim=FALSE
  geom_boxplot(outlier.shape = NA, width=violin_boxPlotWidth) +
  #geom_segment(data = target, aes(x=x, xend=xend, y=value, yend=value), color = "black", linewidth=1, alpha = 0.5, linetype = "dotted") +
  #annotate("text", x = Inf, y = 0.4, label = "RED II revised", angle=90, vjust = -0.5, hjust = 0.075, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  #annotate("text", x = Inf, y = 0.45, label = "REPowerEU", angle=90, vjust = 1.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_hline(yintercept=histValue_2023 %>% pull(value), color = "black", linewidth=1, alpha = 0.5, linetype = "dashed") +
  annotate("text", x = Inf, y = histValue_2023 %>% pull(value), label = "2023", angle=90, vjust = -0.5, hjust = 0, color = "black", alpha = 0.5, size = 0.8*geomText.size) +
  geom_point(data=mainScenarioValue, shape=4, size = 2, stroke = 1.5, alpha = 0.6, color = "red") +
  facet_grid(.~target, scales="free") +
  coord_flip(clip = 'off') +
  scale_y_continuous(breaks=seq(0, 1, 0.2), labels = scales::percent) +
  scale_x_reverse(breaks=c(2030,2040,2050), labels = c(2030,2040,2050), minor_breaks = NULL) +
  expand_limits(y=c(0,1)) + 
  theme_doubleColumn() +
  theme(axis.title.y=element_blank(),
        strip.text.y = element_blank(),
        panel.spacing = unit(4, "lines"),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = textSize.double, hjust = 0, margin=margin(l=0, b=20)),
        #strip.text.x = element_text(size = textSize.double, hjust = 1, margin=margin(l=0, b=20)),
        strip.clip = "off") 

ggsave(paste0(outChartsFolder,"/svg/wind and solar electricity share.svg"), g[["milestone"]][["Wind & Solar Share in Electricity extended"]], device="svg", width = 4*1.5/4, height = 5*2/4, dpi=100, units = "in")
ggsave(paste0(outChartsFolder,"/png/wind and solar electricity share.png"), g[["milestone"]][["Wind & Solar Share in Electricity extended"]], device="png", width = 4*1.5, height = 5*2/4, dpi=300, units = "in")



```

